---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures",
  out.width = "100%"
)
```

## **demulticoder**: An R package for the simultaneous analysis of multiplexed metabarcodes


### Introduction

The ***demulticoder*** package is a cutadapt and DADA2 wrapper package for metabarcodng analyses. The main commands and outputs are intuitive and comprehensive, which helps to account for the complex and iterative nature of metabarcoding analyses. 

Here is a brief schematic of the general workflow:  

```{r workflow-diagram, echo=FALSE, out.width="35%", out.height="35%", fig.align="left"}
knitr::include_graphics("man/figures/Figure1.svg")
```
  
### Key Features

* It automates the use of DADA2 to analyze data derived from multiple metabarcodes.  
* It reduces the number of manual input steps  
* Handles analysis of two metabarcodes multiplexed into the same sequencing batch    
* Analyze different types of metabarcodes simultaneously  
* Reproducible workflows for oomycetes
* Supported metabarcodes: 16S rDNA, ITS1, *rps10*, and up to two additional metabarcodes 

### Installation  

**Dependencies**: 

* DADA2 (TODO-add information on installation)

To install the development version of package: 
```{r, eval=FALSE, echo=TRUE}
library("devtools")
library("DADA2")

devtools::install_github("grunwaldlab/demulticoder")
library("demulticoder")
```

### Quick start  

**1. Set-up input directory and files**  

To demonstrate how to use the package, we have a small test dataset that comes loaded with the package. 
This dataset will be used in the workflow example below.

Already loaded in the test dataset directory are the following files: 

* **PE short read amplicon data**
  * Files: S1_R1.fastq.gz, S1_R2.fastq.gz, S2_R1.fastq.gz, S2_R1.fastq.gz  
  * The files must end in either *R1.fastq.gz* , or *R2.fastq.gz* and each sample must have both R1 and R2 files.
* [**metadata.csv**](https://github.com/grunwaldlab/demulticoder/blob/main/inst/extdata/metadata.csv)
  * New row for each unique sample
  * Samples entered twice if samples contain two pooled metabarcodes, as in the test data template
* [**primerinfo_params.csv**](https://github.com/grunwaldlab/demulticoder/blob/main/inst/extdata/primerinfo_params.csv)
  * New row for each unique barcode and associated primer sequence
  * Optional cutadapt and DADA2 parameters
* **Taxonomy databases**  
  * UNITE fungal database (abridged version)
  * oomyceteDB 
  
See [**Documentation**](https://grunwaldlab.github.io/demulticoder/articles/Documentation.html) for how to format databases and input files. 

For more details on each step, check out the [**Getting Started**](https://grunwaldlab.github.io/demulticoder/articles/Getting_Started.html) tab on the package website

**2. Prepare reads** 
```{r, eval=FALSE, echo=TRUE}
output<-prepare_reads(
  data_directory = system.file("extdata", package = "demulticoder"), # This allows us to use the test directory located within the package
  output_directory = "<OUTDIR>") # Change to you preferred location on your local computer (Example: "~/demulticoder_test")
```

**3. Cut and trim reads** 
User must install cutadapt on their local machine and append the path to the executable.
```{r, eval=FALSE, echo=TRUE}
cut_trim(
  output,
  cutadapt_path="<CUTADAPTPATH>") # Change to the location on your computer. (Example: "/usr/bin/cutadapt")
```
 
  
**4. Make ASV abundance matrix** 
```{r, eval=FALSE, echo=TRUE}
make_asv_abund_matrix(
  output)
```
  
  
**5. Assign taxonomy** 
```{r, eval=FALSE, echo=TRUE}
assign_tax(
  output,
  asv_abund_matrix)
```

  
**6. Convert ASV matrix to taxmap and phyloseq objects** 
```{r, eval=FALSE, echo=TRUE}
objs<-convert_asv_matrix_to_objs(output)
```


### Check out the associated Github repository to view source code  

For more information, to see source code, or submit issue, check out:  
https://github.com/grunwaldlab/demulticoder/


### Citation

The package was developed by Martha Sudermann, Zachary Foster, Samantha Dawson, Hung Phan, Jeff Chang, and Niklaus Grünwald

Stay tuned for associated manuscript. 

### Acknowledgements

The demulticoder logo was created with https://BioRender.com/