---
title: "A comparison of standard workflow vs. demulticoder analysis outputs"
output: 
  rmarkdown::html_vignette:
    fig_path: "man/figures"
vignette: >
  %\VignetteIndexEntry{Multiplexed and pooled barcode examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(phyloseq); packageVersion("phyloseq")
library(ggplot2); packageVersion("ggplot2")
library(readr); packageVersion("readr")
library(tidyr); packageVersion("tidyr")
library(purrr); packageVersion("purrr")
library(furrr); packageVersion("furrr")
library(dplyr); packageVersion("dplyr")
library(stringr); packageVersion("stringr")
library(forcats); packageVersion("forcats")
library(metacoder); packageVersion("metacoder")
library(data.table); packageVersion("data.table")
library(decontam); packageVersion("decontam")
library(Biostrings); packageVersion("Biostrings")
library(magick); packageVersion("magick")
library(vegan); packageVersion("vegan")
library(pdftools);packageVersion("pdftools")
library(vegan); packageVersion("vegan")
library(grid)
library(gridExtra)
```

## More in-depth comparison of standard workflow outputs vs. demulticoder analysis outputs for a multiplexed and pooled pooled barcode dataset

This analysis accompanies the rhododendron rhizobiome analyses described in the associated manuscript. 

Briefly, TODO

### Input metadata and primerinfo_params files are in data folder
```{r setup, include=FALSE}
#TODO-at some point switch to loading github repo version
devtools::install_github("grunwaldlab/demulticoder", force=TRUE)
library("demulticoder")
```

### Load demulticoder outputs
```{r load demulticoder outputs}
seed <- 1
set.seed(seed)

abundance_with_inf <- read.csv("~/benchmark_demulticoder/demulticoder/data/abundance_with_infection_data.csv") 

metadata_path <- file.path("~/benchmark_demulticoder/demulticoder/data/metadata_with_infection_data.csv")
metadata <- readr::read_csv(metadata_path)
print(metadata)
```

### Load standard workflow outputs
```{r load standard workflow outputs}
abundance_sw_inf <- read.csv("~/benchmark_demulticoder/standard_workflow/combined_data/abundance_with_infection_data.csv")
#Let's also load metadata file-consistent with demulticoder analysis
metadata_path_sf <- file.path("~/benchmark_demulticoder/standard_workflow/combined_data/metadata_with_infection_data.csv")
metadata_sf <- read_csv(metadata_path_sf)
print(metadata_sf)
```

### Generate summary of read counts throughout the workflow for demlticoder specific analysis
```{r get demulticoder read stats}
#Fourth, let's track reads
track_reads_demulticoder_its<-read.csv("~/benchmark_demulticoder/demulticoder/data/track_reads_its.csv", row.names = 1)
track_reads_demulticoder_rps10<-read.csv("~/benchmark_demulticoder/demulticoder/data/track_reads_rps10.csv", row.names = 1)
track_reads_sw_its<-read.csv("~/benchmark_demulticoder/standard_workflow/its/data/its_trackreads.out", row.names = 1)
track_reads_standard_wf_its <- read.table("~/benchmark_demulticoder/standard_workflow/its/data/its_trackreads.out", sep="\t", stringsAsFactors = F, header=T)
track_reads_standard_wf_rps10 <- read.table("~/benchmark_demulticoder/standard_workflow/rps10/data/rps10_trackreads.out", sep="\t", stringsAsFactors = F, header=T)
```

### Now, let's compare read counts across workflows for rps10 amplicons
```{r compare reads-rps10}
rps10_summary_reads_demulticoder<-summary(track_reads_demulticoder_rps10)
print(rps10_summary_reads_demulticoder)

rps10_summary_reads_sw<-summary(track_reads_standard_wf_rps10)
print(rps10_summary_reads_sw)
```

### Now, let's compare read counts across workflows for ITS amplicons
```{r compare reads-rps10}
its_summary_reads_demulticoder<-summary(track_reads_demulticoder_its)
print(its_summary_reads_demulticoder)

its_summary_reads_sw<-summary(track_reads_standard_wf_its)
print(its_summary_reads_sw)
```

### Let's convert our two matrices to taxmap objects
Let's convert our new matrix and associated metadata table to a taxmap object for demulticoder analysis
```{r, Convert matrix created with demulticoder workflow to taxmap obj}
metadata <- filter(metadata, ! is_ambiguous)
abundance_with_inf <- filter(abundance_with_inf, ! is_inoculum)

#Filter out mock community samples and also any mock community samples as well
metadata <- filter(metadata, sample_type != "Mock community" & sample_type != "Negative control")

obj <- parse_tax_data(abundance_with_inf, class_cols = 'dada2_tax', class_sep = ';',
                      class_regex = '^(.+)--(.+)--(.+)$',
                      class_key = c(taxon = 'taxon_name', boot = 'info', rank = 'taxon_rank'))
names(obj$data) <- c('abund', 'score')
obj <- transmute_obs(obj, 'score', sequence = sequence[input_index], boot = boot, rank = rank)


# For diversity calculations, we'll use proportions of read depth
# We'll set any proportion to 0 that is less than the inverse of the read count of the non-control sample with the fewest reads.
# This should account for unequal sample read depth without the randomness of rarefaction.

metadata$raw_count <- colSums(obj$data$abund[, metadata$sample_name])
lowest_count <- min(metadata$raw_count[! is.na(metadata$organism)])
lowest_count
obj$data$prop <- calc_obs_props(obj, data = 'abund', cols = metadata$sample_name)
obj$data$prop <- zero_low_counts(obj, data = 'prop', min_count = 1 / lowest_count, cols = metadata$sample_name)
obj$data$prop

obj$data$prop[metadata$sample_name] <- map(metadata$sample_name, function(id) {
  out <- obj$data$prop[[id]]
  out[is.na(out) | is.nan(out)] <- 0
  out
})

obj$data$prop

#Let's modify the metadata sheet
metadata <- metadata %>%
  mutate(organism = fct_relevel(ordered(organism), "Control", "Cin", "Plu", "Cry"))
```

Let's convert our new matrix and associated metadata table to a taxmap object for output of standard workflow analysis
```{r, Convert matrix created with demulticoder workflow to taxmap obj}
metadata_sf <- filter(metadata_sf, ! is_ambiguous)
abundance_sw_inf <- filter(abundance_sw_inf, ! is_inoculum)

#Filter out mock community samples and also any mock community samples as well
metadata_sf <- filter(metadata_sf, sample_type != "Mock community" & sample_type != "Negative control")

obj_sf <- parse_tax_data(abundance_sw_inf, class_cols = 'dada2_tax', class_sep = ';',
                      class_regex = '^(.+)--(.+)--(.+)$',
                      class_key = c(taxon = 'taxon_name', boot = 'info', rank = 'taxon_rank'))
names(obj_sf$data) <- c('abund', 'score')
obj_sf <- transmute_obs(obj_sf, 'score', sequence = sequence[input_index], boot = boot, rank = rank)


# For diversity calculations, we'll use proportions of read depth
# We'll set any proportion to 0 that is less than the inverse of the read count of the non-control sample with the fewest reads.
# This should account for unequal sample read depth without the randomness of rarefaction.

metadata_sf$raw_count <- colSums(obj_sf$data$abund[, metadata_sf$sample_name])
lowest_count <- min(metadata_sf$raw_count[! is.na(metadata_sf$organism)])
lowest_count
obj_sf$data$prop <- calc_obs_props(obj_sf, data = 'abund', cols = metadata_sf$sample_name)
obj_sf$data$prop <- zero_low_counts(obj_sf, data = 'prop', min_count = 1 / lowest_count, cols = metadata_sf$sample_name)
obj_sf$data$prop

obj_sf$data$prop[metadata_sf$sample_name] <- map(metadata_sf$sample_name, function(id) {
  out <- obj_sf$data$prop[[id]]
  out[is.na(out) | is.nan(out)] <- 0
  out
})

obj_sf$data$prop

#Let's modify the metadata_sf sheet
metadata_sf <- metadata_sf %>%
  mutate(organism = fct_relevel(ordered(organism), "Control", "Cin", "Plu", "Cry"))
```

### Let's examine alpha diversity within pots
Demulticoder outputs
```{r, alpha diversity demulticoder}
abund_table <- obj$data$abund[metadata$sample_name]
metadata$richness <- vegan::specnumber(abund_table, MARGIN = 2)
metadata$shannon <- vegan::diversity(abund_table, MARGIN = 2, index = "shannon")
metadata$invsimpson <- vegan::diversity(abund_table, MARGIN = 2, index = "invsimpson")

plotted_factors <- c('Organism' = 'organism', 'Flooded' = 'flooded', 'Pathogen Concentration' = 'path_conc', 'Trial' = 'experiment')

# Reformat data for plotting
alpha_plot_data <- plotted_factors %>%
  map2_dfr(names(plotted_factors), 
           function(factor, factor_name) {
             out <- metadata
             out$factor <- factor_name
             out$value <- as.character(metadata[[factor]])
             return(out)
           }) %>%
  mutate(path_conc = factor(path_conc, 
                            levels = sort(unique(path_conc)), 
                            labels = paste(sort(unique(path_conc)), 'CFU/g'), 
                            ordered = TRUE)) %>%
  filter(sample_type == 'Sample') %>%
  select(sample_name, factor, value, invsimpson) %>%
  tidyr::gather(key = "index", value = "diversity", -sample_name, -factor, -value) %>%
  mutate(value = forcats::fct_relevel(ordered(value), "Control", "Cin", "Plu", "Cry"))

# ANOVA and Tukey's HSD
anova_and_hsd <- function(x) {
  anova_result <- aov(diversity ~ value, x)
  tukey_result <- agricolae::HSD.test(anova_result, "value", group = TRUE)
  group_data <- tukey_result$groups[order(rownames(tukey_result$groups)),]
  group_key <- setNames(group_data$groups, rownames(group_data))
  group_key[as.character(x$value)]
}
alpha_plot_data$group <- unlist(map(split(alpha_plot_data, alpha_plot_data$factor)[unique(alpha_plot_data$factor)], anova_and_hsd))


alpha_subplot <- ggplot(alpha_plot_data, aes(x = value, y = diversity)) +
  geom_boxplot() +
  geom_text(aes(x = value,
                y = max(diversity) + 2,
                label = group),
            col = 'black',
            size = 5) +
  facet_grid( ~ factor, scales = "free") +
  labs(x = NULL, y = 'Diversity (Inverse Simpson)') +
  guides(color = "none") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom")

alpha_subplot
```

Standard workflow
```{r, alpha diversity demulticoder}
abund_table_sf <- obj_sf$data$abund[metadata_sf$sample_name]
metadata_sf$richness <- vegan::specnumber(abund_table_sf, MARGIN = 2)
metadata_sf$shannon <- vegan::diversity(abund_table_sf, MARGIN = 2, index = "shannon")
metadata_sf$invsimpson <- vegan::diversity(abund_table_sf, MARGIN = 2, index = "invsimpson")

plotted_factors <- c('Organism' = 'organism', 'Flooded' = 'flooded', 'Pathogen Concentration' = 'path_conc', 'Trial' = 'experiment')

# Reformat data for plotting
alpha_plot_data_sf <- plotted_factors %>%
  map2_dfr(names(plotted_factors), 
           function(factor, factor_name) {
             out <- metadata_sf
             out$factor <- factor_name
             out$value <- as.character(metadata_sf[[factor]])
             return(out)
           }) %>%
  mutate(path_conc = factor(path_conc, 
                            levels = sort(unique(path_conc)), 
                            labels = paste(sort(unique(path_conc)), 'CFU/g'), 
                            ordered = TRUE)) %>%
  filter(sample_type == 'Sample') %>%
  select(sample_name, factor, value, invsimpson) %>%
  tidyr::gather(key = "index", value = "diversity", -sample_name, -factor, -value) %>%
  mutate(value = forcats::fct_relevel(ordered(value), "Control", "Cin", "Plu", "Cry"))

# ANOVA and Tukey's HSD
anova_and_hsd_sf <- function(x) {
  anova_result <- aov(diversity ~ value, x)
  tukey_result <- agricolae::HSD.test(anova_result, "value", group = TRUE)
  group_data <- tukey_result$groups[order(rownames(tukey_result$groups)),]
  group_key <- setNames(group_data$groups, rownames(group_data))
  group_key[as.character(x$value)]
}
alpha_plot_data_sf$group <- unlist(map(split(alpha_plot_data_sf, alpha_plot_data_sf$factor)[unique(alpha_plot_data_sf$factor)], anova_and_hsd_sf))


alpha_subplot_sf <- ggplot(alpha_plot_data_sf, aes(x = value, y = diversity)) +
  geom_boxplot() +
  geom_text(aes(x = value,
                y = max(diversity) + 2,
                label = group),
            col = 'black',
            size = 5) +
  facet_grid( ~ factor, scales = "free") +
  labs(x = NULL, y = 'Diversity (Inverse Simpson)') +
  guides(color = "none") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom")

alpha_subplot_sf
```

### Let's examine beta diversity within pots
Demulticoder outputs
```{r, beta diversity demulticoder}
set.seed(1)
prob_table <- obj$data$prop[metadata$sample_name]
nmds_plot_data <- function(prob_table) {
  metadata <- metadata[metadata$sample_name %in% colnames(prob_table), ]
  set.seed(1)
  nmds_results <- vegan::metaMDS(t(prob_table), trymax = 1000, k = 2, trace = 0)
  nmds_data <- nmds_results$points %>%
    as_tibble() %>%
    bind_cols(metadata)
  names(nmds_data)[1:2] <- paste0("NMDS", 1:2)
  return(nmds_data)
}

nmds_data <- nmds_plot_data(prob_table[! is.na(metadata$organism) & metadata$valid_inoc])

nmds_factors <- c(Flooded = 'flooded', Organism = 'organism', 'Pathogen CFU/g' = 'path_conc', 'Trial' = 'experiment')

make_one_plot <- function(factor, name) {
  nmds_data %>%
    mutate(factor = as.character(nmds_data[[factor]]),
           NMDS1 = scales::rescale(NMDS1),
           NMDS2 = scales::rescale(NMDS2)) %>%
    mutate(factor = fct_relevel(ordered(factor), "Control", "Cin", "Plu", "Cry")) %>%
    ggplot(aes_string(x = "NMDS1", y = "NMDS2", color = "factor", label = "sample_name")) +
    # geom_label(size = 2) +
    geom_point() +
    coord_fixed() +
    viridis::scale_color_viridis(discrete = TRUE, end = .9) +
    labs(color = NULL, x = NULL, y = NULL) +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text = element_blank(),
          axis.title = element_text(size = 7), 
          axis.ticks = element_blank(),
          plot.margin = unit(rep(0.04, 4), "cm"),
          # panel.background = element_rect(fill = 'transparent', colour = NA),
          # plot.background = element_rect(fill = "white", colour = NA),
          legend.position = "bottom",
          legend.text = element_text(size = 10),
          legend.key.height = unit(0.5, 'cm'),
          legend.key.width = unit(0.2, 'cm'))
}

nmds_subplots <- map2(nmds_factors, names(nmds_factors), make_one_plot)
nmds_plot <- ggpubr::ggarrange(plotlist = c(list(ggplot() + theme_void()), nmds_subplots),
                       nrow = 1, widths = c(0.15, 1, 1, 1, 1))

nmds_plot
```

Standard workflow
```{r, beta diversity demulticoder}
set.seed(1)
prob_table <- obj_sf$data$prop[metadata$sample_name]
nmds_plot_sf_data_sf <- function(prob_table) {
  metadata <- metadata[metadata$sample_name %in% colnames(prob_table), ]
  set.seed(1)
  nmds_results <- vegan::metaMDS(t(prob_table), trymax = 1000, k = 2, trace = 0)
  nmds_data_sf <- nmds_results$points %>%
    as_tibble() %>%
    bind_cols(metadata)
  names(nmds_data_sf)[1:2] <- paste0("NMDS", 1:2)
  return(nmds_data_sf)
}

nmds_data_sf <- nmds_plot_sf_data_sf(prob_table[! is.na(metadata$organism) & metadata$valid_inoc])

nmds_factors <- c(Flooded = 'flooded', Organism = 'organism', 'Pathogen CFU/g' = 'path_conc', 'Trial' = 'experiment')

make_one_plot <- function(factor, name) {
  nmds_data_sf %>%
    mutate(factor = as.character(nmds_data_sf[[factor]]),
           NMDS1 = scales::rescale(NMDS1),
           NMDS2 = scales::rescale(NMDS2)) %>%
    mutate(factor = fct_relevel(ordered(factor), "Control", "Cin", "Plu", "Cry")) %>%
    ggplot(aes_string(x = "NMDS1", y = "NMDS2", color = "factor", label = "sample_name")) +
    # geom_label(size = 2) +
    geom_point() +
    coord_fixed() +
    viridis::scale_color_viridis(discrete = TRUE, end = .9) +
    labs(color = NULL, x = NULL, y = NULL) +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text = element_blank(),
          axis.title = element_text(size = 7), 
          axis.ticks = element_blank(),
          plot.margin = unit(rep(0.04, 4), "cm"),
          # panel.background = element_rect(fill = 'transparent', colour = NA),
          # plot.background = element_rect(fill = "white", colour = NA),
          legend.position = "bottom",
          legend.text = element_text(size = 10),
          legend.key.height = unit(0.5, 'cm'),
          legend.key.width = unit(0.2, 'cm'))
}

nmds_subplots <- map2(nmds_factors, names(nmds_factors), make_one_plot)
nmds_plot_sf <- ggpubr::ggarrange(plotlist = c(list(ggplot() + theme_void()), nmds_subplots),
                       nrow = 1, widths = c(0.15, 1, 1, 1, 1))

 nmds_plot_sf
```

### Let's do a quick differential abundance analysis with DESeq2 and look at differences when inoculum concentration is altered 
Demulticoder outputs
```{r diff heat tree setup demulticoder}
#Make sure our factors are input properly as factors
metadata$path_conc <- as.factor(metadata$path_conc)
metadata$path_conc<- factor(metadata$path_conc, levels = c("0", "1", "100"))

metadata$organism <- as.factor(metadata$organism)
metadata$organism<- factor(metadata$organism, levels = c("Control", "Cin", "Cry", "Plu"))

metadata$flooded <- as.factor(metadata$flooded)
metadata$flooded <- factor(metadata$flooded, levels = c("TRUE", "FALSE"))

metadata$experiment <- as.factor(metadata$experiment)
metadata$experiment<- factor(metadata$experiment, levels = c("1", "2"))

abundance_with_inf$is_low_abund <- rowSums(abundance_with_inf[, metadata$sample_name]) < 5
obj2 <- parse_tax_data(abundance_with_inf, class_cols = 'dada2_tax', class_sep = ';',
                      class_regex = '^(.+)--(.+)--(.+)$',
                      class_key = c(taxon = 'taxon_name', boot = 'info', rank = 'taxon_rank'))
names(obj2$data) <- c('abund', 'score')
obj2 <- transmute_obs(obj2, 'score', sequence = sequence[input_index], boot = boot, rank = rank)
print(obj2)

obj2$data$tax_abund <- calc_taxon_abund(obj2, data = "abund", cols = metadata$sample_name)

obj2$data$abund_prop <- calc_obs_props(obj2, data = "abund", cols = metadata$sample_name, groups = rep("tax_prop", nrow(metadata)))
obj2$data$tax_prop <- calc_taxon_abund(obj2, data = "abund_prop", cols = "tax_prop")
obj2$data$abund_prop <- NULL

obj_subset <- obj2 %>%
  filter_taxa(taxon_ranks == "Species", supertaxa = TRUE, reassign_obs = c(tax_abund = FALSE)) 
min_bootstrap <- 0

obj_subset$data$score$boot <- as.numeric(obj_subset$data$score$boot)
max_boot <- obj_subset$data$score %>%
  group_by(taxon_id) %>%
  summarise(max = max(boot))
max_boot <- setNames(max_boot$max, max_boot$taxon_id)
obj_subset <- filter_taxa(obj_subset, max_boot[taxon_ids] >= min_bootstrap | taxon_ranks %in% c("ASV", "Reference"), reassign_obs = c(abund = TRUE, score = FALSE))
```

Standard workflow outputs
```{r diff heat tree setup standard workflow}
#Make sure our factors are input properly as factors
metadata_sf$path_conc <- as.factor(metadata_sf$path_conc)
metadata_sf$path_conc<- factor(metadata_sf$path_conc, levels = c("0", "1", "100"))

metadata_sf$organism <- as.factor(metadata_sf$organism)
metadata_sf$organism<- factor(metadata_sf$organism, levels = c("Control", "Cin", "Cry", "Plu"))

metadata_sf$flooded <- as.factor(metadata_sf$flooded)
metadata_sf$flooded <- factor(metadata_sf$flooded, levels = c("TRUE", "FALSE"))

metadata_sf$experiment <- as.factor(metadata_sf$experiment)
metadata_sf$experiment<- factor(metadata_sf$experiment, levels = c("1", "2"))

abundance_sw_inf$is_low_abund <- rowSums(abundance_sw_inf[, metadata_sf$sample_name]) < 5
obj2_sf <- parse_tax_data(abundance_sw_inf, class_cols = 'dada2_tax', class_sep = ';',
                      class_regex = '^(.+)--(.+)--(.+)$',
                      class_key = c(taxon = 'taxon_name', boot = 'info', rank = 'taxon_rank'))
names(obj2_sf$data) <- c('abund', 'score')
obj2_sf <- transmute_obs(obj2_sf, 'score', sequence = sequence[input_index], boot = boot, rank = rank)
print(obj2_sf)

obj2_sf$data$tax_abund <- calc_taxon_abund(obj2_sf, data = "abund", cols = metadata_sf$sample_name)

obj2_sf$data$abund_prop <- calc_obs_props(obj2_sf, data = "abund", cols = metadata_sf$sample_name, groups = rep("tax_prop", nrow(metadata_sf)))
obj2_sf$data$tax_prop <- calc_taxon_abund(obj2_sf, data = "abund_prop", cols = "tax_prop")
obj2_sf$data$abund_prop <- NULL

obj_subset_sf <- obj2_sf %>%
  filter_taxa(taxon_ranks == "Species", supertaxa = TRUE, reassign_obs = c(tax_abund = FALSE)) 
min_bootstrap <- 0

obj_subset_sf$data$score$boot <- as.numeric(obj_subset_sf$data$score$boot)
max_boot <- obj_subset_sf$data$score %>%
  group_by(taxon_id) %>%
  summarise(max = max(boot))
max_boot <- setNames(max_boot$max, max_boot$taxon_id)
obj_subset_sf <- filter_taxa(obj_subset_sf, max_boot[taxon_ids] >= min_bootstrap | taxon_ranks %in% c("ASV", "Reference"), reassign_obs = c(abund = TRUE, score = FALSE))
```

### Let's make differential abundance plots for our combined dataset-first let's look at diff. abundance of taxa by whether pots were flooded or not

Demulticoder analysis
```{r abundance plots-inoculum, fig.height=20, fig.width=20, message=FALSE, echo=FALSE}
# Calculate difference between groups
obj_subset$data$diff_table_flooded <- calc_diff_abund_deseq2(obj_subset, data = "tax_abund",
                                                         cols = metadata$sample_name[!is.na(metadata$flooded)],
                                                         groups = metadata$flooded[!is.na(metadata$flooded)])
significant_taxon_ids <- unique(
  obj_subset$data$diff_table_flooded$taxon_id[! is.na(obj_subset$data$diff_table_flooded$padj) & obj_subset$data$diff_table_flooded$padj < 0.05]
)

range(obj_subset$data$diff_table_flooded$padj, finite = TRUE) 
range(obj_subset$data$diff_table_flooded$log2FoldChange, finite = TRUE) 

set.seed = 1000
# Plot results (might take a few minutes)
obj_subset %>%
  filter_taxa(taxon_ids %in% significant_taxon_ids, supertaxa = TRUE, reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "_sp$"), reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "incertae_sedis", ignore.case = TRUE), reassign_obs = FALSE) %>%
  remove_redundant_names() %>%
  heat_tree_matrix(data = "diff_table_flooded",
                   node_size = tax_prop,
                   node_label = taxon_names,
                   node_color = ifelse(is.na(padj) | padj > 0.05, 0, log2FoldChange),
                   node_color_range = diverging_palette(),
                   node_color_trans = "linear",
                   node_color_interval = c(-7.5, 7.5),
                   edge_color_interval = c(-7.5, 7.5),
                   node_size_range = c(.005, .035),
                   node_label_size_range = c(.015, .025),
                   node_size_trans = "log10",
                   node_size_axis_label = "Mean read proportion",
                   node_color_axis_label = "Log2 fold change",
                   node_size_digits = 1,
                   seed=1000,
                   key_size = 0.5)

obj_subset
```

Standard workflow
```{r, diff abundance standard workflow}
# Calculate difference between groups
obj_subset_sf$data$diff_table_flooded <- calc_diff_abund_deseq2(obj_subset_sf, data = "tax_abund",
                                                         cols = metadata$sample_name[!is.na(metadata$flooded)],
                                                         groups = metadata$flooded[!is.na(metadata$flooded)])
significant_taxon_ids <- unique(
  obj_subset_sf$data$diff_table_flooded$taxon_id[! is.na(obj_subset_sf$data$diff_table_flooded$padj) & obj_subset_sf$data$diff_table_flooded$padj < 0.05]
)

range(obj_subset_sf$data$diff_table_flooded$padj, finite = TRUE) 
range(obj_subset_sf$data$diff_table_flooded$log2FoldChange, finite = TRUE) 

set.seed = 1000
# Plot results (might take a few minutes)
obj_subset_sf %>%
  filter_taxa(taxon_ids %in% significant_taxon_ids, supertaxa = TRUE, reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "_sp$"), reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "incertae_sedis", ignore.case = TRUE), reassign_obs = FALSE) %>%
  remove_redundant_names() %>%
  heat_tree_matrix(data = "diff_table_flooded",
                   node_size = tax_prop,
                   node_label = taxon_names,
                   node_color = ifelse(is.na(padj) | padj > 0.05, 0, log2FoldChange),
                   node_color_range = diverging_palette(),
                   node_color_trans = "linear",
                   node_color_interval = c(-7.5, 7.5),
                   edge_color_interval = c(-7.5, 7.5),
                   node_size_range = c(.005, .035),
                   node_label_size_range = c(.015, .025),
                   node_size_trans = "log10",
                   node_size_axis_label = "Mean read proportion",
                   node_color_axis_label = "Log2 fold change",
                   node_size_digits = 1,
                   seed=1000,
                   key_size = 0.5)

obj_subset_sf
```


### Let's make differential abundance plots for our combined dataset-first let's look at diff. abundance of taxa by which pathogen was used as inoculum

Demulticoder analysis
```{r abundance plots-inoculum, fig.height=20, fig.width=20, message=FALSE, echo=FALSE}
# Calculate difference between groups
obj_subset$data$diff_table_org <- calc_diff_abund_deseq2(obj_subset, data = "tax_abund",
                                                         cols = metadata$sample_name[!is.na(metadata$organism)],
                                                         groups = metadata$organism[!is.na(metadata$organism)])
significant_taxon_ids <- unique(
  obj_subset$data$diff_table_org$taxon_id[! is.na(obj_subset$data$diff_table_org$padj) & obj_subset$data$diff_table_org$padj < 0.05]
)

range(obj_subset$data$diff_table_org$padj, finite = TRUE) 
range(obj_subset$data$diff_table_org$log2FoldChange, finite = TRUE) 

set.seed = 1000
# Plot results (might take a few minutes)
obj_subset %>%
  filter_taxa(taxon_ids %in% significant_taxon_ids, supertaxa = TRUE, reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "_sp$"), reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "incertae_sedis", ignore.case = TRUE), reassign_obs = FALSE) %>%
  remove_redundant_names() %>%
  heat_tree_matrix(data = "diff_table_org",
                   node_size = tax_prop,
                   node_label = taxon_names,
                   node_color = ifelse(is.na(padj) | padj > 0.05, 0, log2FoldChange),
                   node_color_range = diverging_palette(),
                   node_color_trans = "linear",
                   node_color_interval = c(-7.5, 7.5),
                   edge_color_interval = c(-7.5, 7.5),
                   node_size_range = c(.005, .035),
                   node_label_size_range = c(.015, .025),
                   node_size_trans = "log10",
                   node_size_axis_label = "Mean read proportion",
                   node_color_axis_label = "Log2 fold change",
                   node_size_digits = 1,
                   seed=1000,
                   key_size = 0.5)

obj_subset
```

Standard workflow
```{r, diff abundance standard workflow}
# Calculate difference between groups
obj_subset_sf$data$diff_table_org <- calc_diff_abund_deseq2(obj_subset_sf, data = "tax_abund",
                                                         cols = metadata$sample_name[!is.na(metadata$organism)],
                                                         groups = metadata$organism[!is.na(metadata$organism)])
significant_taxon_ids <- unique(
  obj_subset_sf$data$diff_table_org$taxon_id[! is.na(obj_subset_sf$data$diff_table_org$padj) & obj_subset_sf$data$diff_table_org$padj < 0.05]
)

range(obj_subset_sf$data$diff_table_org$padj, finite = TRUE) 
range(obj_subset_sf$data$diff_table_org$log2FoldChange, finite = TRUE) 

set.seed = 1000
# Plot results (might take a few minutes)
obj_subset_sf %>%
  filter_taxa(taxon_ids %in% significant_taxon_ids, supertaxa = TRUE, reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "_sp$"), reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "incertae_sedis", ignore.case = TRUE), reassign_obs = FALSE) %>%
  remove_redundant_names() %>%
  heat_tree_matrix(data = "diff_table_org",
                   node_size = tax_prop,
                   node_label = taxon_names,
                   node_color = ifelse(is.na(padj) | padj > 0.05, 0, log2FoldChange),
                   node_color_range = diverging_palette(),
                   node_color_trans = "linear",
                   node_color_interval = c(-7.5, 7.5),
                   edge_color_interval = c(-7.5, 7.5),
                   node_size_range = c(.005, .035),
                   node_label_size_range = c(.015, .025),
                   node_size_trans = "log10",
                   node_size_axis_label = "Mean read proportion",
                   node_color_axis_label = "Log2 fold change",
                   node_size_digits = 1,
                   seed=1000,
                   key_size = 0.5)

obj_subset_sf
```


### Let's examine community composition more broadly
Demulticoder outputs
```{r, community composition demulticoder}
obj_subset <- obj2 %>%
  filter_taxa(taxon_ranks == "Species", supertaxa = TRUE, reassign_obs = c(tax_abund = FALSE)) 
min_bootstrap <- 60

obj_subset$data$score$boot <- as.numeric(obj_subset$data$score$boot)
max_boot <- obj_subset$data$score %>%
  group_by(taxon_id) %>%
  summarise(max = max(boot))
max_boot <- setNames(max_boot$max, max_boot$taxon_id)
obj_subset <- filter_taxa(obj_subset, max_boot[taxon_ids] >= min_bootstrap | taxon_ranks %in% c("ASV", "Reference"), reassign_obs = c(abund = TRUE, score = FALSE))

obj_subset$data$asv_prop <- calc_obs_props(obj_subset, 'abund', cols = metadata$sample_name)
obj_subset$data$tax_abund <- calc_taxon_abund(obj_subset, 'abund', cols = metadata$sample_name)
obj_subset$data$tax_prop <- calc_taxon_abund(obj_subset, 'asv_prop', cols = metadata$sample_name)
obj_subset$data$tax_data <- calc_n_samples(obj_subset, 'tax_prop', cols = metadata$sample_name[metadata$sample_type == 'Sample'])
obj_subset$data$tax_data$mean_prop <- rowMeans(obj_subset$data$tax_prop[, metadata$sample_name])

# Calculate the relative standard deviation for each taxon as a measure of how consistently it was found.
rsd <- function(x, na.rm = FALSE) {sd(x, na.rm = na.rm) / mean(x, na.rm = na.rm)}
obj_subset$data$tax_data$rel_stand_dev <- map_dbl(1:nrow(obj_subset$data$tax_abund), function(i) {
  rsd(unlist(obj_subset$data$tax_abund[i, metadata$sample_name]), na.rm = TRUE)
})
obj_subset

set.seed(1000)
obj_subset %>%
  filter_taxa(! is_stem) %>%
  filter_taxa(n_samples >= 10, supertaxa = TRUE, reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "_sp$"), reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "incertae_sedis", ignore.case = TRUE), reassign_obs = FALSE) %>%
  remove_redundant_names() %>%
  heat_tree(node_size = mean_prop,
            edge_size = n_samples,
            node_color = ifelse(is.na(rel_stand_dev), 0, rel_stand_dev),
            #node_color_range = c("#aaaaaa", "#8da0cb", "#66c2a5", "#a6d854", "#fc8d62", "red"),
            node_label = taxon_names,
            node_size_range = c(0.008, 0.025),
            node_label_size_range = c(0.012, 0.018),
            edge_label_size_range = c(0.010, 0.013),
            node_size_interval = c(0, 1),
            edge_size_range = c(0.001, 0.008),
            #layout = "da", initial_layout = "re",
            node_color_axis_label = "Relative standard deviation",
            node_size_axis_label = "Mean proportion of reads",
            edge_size_axis_label = "Number of samples",
            node_color_digits = 2,
            node_size_digits = 2,
            edge_color_digits = 2,
            edge_size_digits = 2)

obj_subset
```

Standard workflow
```{r, community composition standard workflow}
obj_subset_sf <- obj2 %>%
  filter_taxa(taxon_ranks == "Species", supertaxa = TRUE, reassign_obs = c(tax_abund = FALSE)) 
min_bootstrap <- 60

obj_subset_sf$data$score$boot <- as.numeric(obj_subset_sf$data$score$boot)
max_boot <- obj_subset_sf$data$score %>%
  group_by(taxon_id) %>%
  summarise(max = max(boot))
max_boot <- setNames(max_boot$max, max_boot$taxon_id)
obj_subset_sf <- filter_taxa(obj_subset_sf, max_boot[taxon_ids] >= min_bootstrap | taxon_ranks %in% c("ASV", "Reference"), reassign_obs = c(abund = TRUE, score = FALSE))

obj_subset_sf$data$asv_prop <- calc_obs_props(obj_subset_sf, 'abund', cols = metadata$sample_name)
obj_subset_sf$data$tax_abund <- calc_taxon_abund(obj_subset_sf, 'abund', cols = metadata$sample_name)
obj_subset_sf$data$tax_prop <- calc_taxon_abund(obj_subset_sf, 'asv_prop', cols = metadata$sample_name)
obj_subset_sf$data$tax_data <- calc_n_samples(obj_subset_sf, 'tax_prop', cols = metadata$sample_name[metadata$sample_type == 'Sample'])
obj_subset_sf$data$tax_data$mean_prop <- rowMeans(obj_subset_sf$data$tax_prop[, metadata$sample_name])

# Calculate the relative standard deviation for each taxon as a measure of how consistently it was found.
rsd <- function(x, na.rm = FALSE) {sd(x, na.rm = na.rm) / mean(x, na.rm = na.rm)}
obj_subset_sf$data$tax_data$rel_stand_dev <- map_dbl(1:nrow(obj_subset_sf$data$tax_abund), function(i) {
  rsd(unlist(obj_subset_sf$data$tax_abund[i, metadata$sample_name]), na.rm = TRUE)
})
obj_subset_sf

set.seed(1000)
obj_subset_sf %>%
  filter_taxa(! is_stem) %>%
  filter_taxa(n_samples >= 10, supertaxa = TRUE, reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "_sp$"), reassign_obs = FALSE) %>%
  filter_taxa(! grepl(x = taxon_names, "incertae_sedis", ignore.case = TRUE), reassign_obs = FALSE) %>%
  remove_redundant_names() %>%
  heat_tree(node_size = mean_prop,
            edge_size = n_samples,
            node_color = ifelse(is.na(rel_stand_dev), 0, rel_stand_dev),
            #node_color_range = c("#aaaaaa", "#8da0cb", "#66c2a5", "#a6d854", "#fc8d62", "red"),
            node_label = taxon_names,
            node_size_range = c(0.008, 0.025),
            node_label_size_range = c(0.012, 0.018),
            edge_label_size_range = c(0.010, 0.013),
            node_size_interval = c(0, 1),
            edge_size_range = c(0.001, 0.008),
            #layout = "da", initial_layout = "re",
            node_color_axis_label = "Relative standard deviation",
            node_size_axis_label = "Mean proportion of reads",
            edge_size_axis_label = "Number of samples",
            node_color_digits = 2,
            node_size_digits = 2,
            edge_color_digits = 2,
            edge_size_digits = 2)
            #aspect_ratio = 1.618)

obj_subset_sf
```

```{r}
sessioninfo::session_info()
```