---
title: "A comparison of standard workflow vs. demulticoder analysis outputs"
output: 
  rmarkdown::html_vignette:
    fig_path: "man/figures"
vignette: >
  %\VignetteIndexEntry{Multiplexed and pooled barcode examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(phyloseq); packageVersion("phyloseq")
library(ggplot2); packageVersion("ggplot2")
library(readr); packageVersion("readr")
library(tidyr); packageVersion("tidyr")
library(purrr); packageVersion("purrr")
library(furrr); packageVersion("furrr")
library(dplyr); packageVersion("dplyr")
library(stringr); packageVersion("stringr")
library(forcats); packageVersion("forcats")
library(metacoder); packageVersion("metacoder")
library(data.table); packageVersion("data.table")
library(decontam); packageVersion("decontam")
library(Biostrings); packageVersion("Biostrings")
library(magick); packageVersion("magick")
library(vegan); packageVersion("vegan")
library(pdftools);packageVersion("pdftools")
library(vegan); packageVersion("vegan")
library(grid)
library(gridExtra)
```

## More in-depth comparison of standard workflow outputs vs. demluticoder analysis outputs for a multiplexed and pooled pooled barcode dataset

This analysis accompanies the analyses described in the associated manuscript. 
Input metadata and primerinfo_params files are in data folder

```{r setup, include=FALSE}
#TODO-at some point switch to loading github repo version
devtools::load_all("~/demulticoder")
library("demulticoder")
```

### Load demulticoder outputs
```{r load demulticoder outputs}
seed <- 1
set.seed(seed)

asv_matrix_rps10<-read.csv("demulticoder/data/final_asv_abundance_matrix_rps10.csv")
asv_matrix_rps10$dada2_tax <- asv_matrix_rps10$dada2_tax <- gsub("Stramenopila", "Eukaryota--100--Domain;Stramenopila", asv_matrix_rps10$dada2_tax)
asv_matrix_rps10 <- asv_matrix_rps10[, -1]
colnames(asv_matrix_rps10)[3:ncol(asv_matrix_rps10)] <- gsub("_rps10$", "", colnames(asv_matrix_rps10)[3:ncol(asv_matrix_rps10)])

asv_matrix_its<-read.csv("demulticoder/data/final_asv_abundance_matrix_its.csv")
asv_matrix_its$dada2_tax <- gsub("Fungi", "Eukaryota--100--Domain;Fungi", asv_matrix_its$dada2_tax)
asv_matrix_its <- asv_matrix_its[, -1]
colnames(asv_matrix_its)[3:ncol(asv_matrix_rps10)] <- gsub("_its$", "", colnames(asv_matrix_its)[3:ncol(asv_matrix_its)])


#Let's combine these matrices
#For easier analysis, we previously combined the two matrices, and appended domain info to each one so we can make one heat tree for combined dataset
sample_cols_its <- setdiff(names(asv_matrix_its), c("sequence", "dada2_tax"))
sample_cols_rps10 <- setdiff(names(asv_matrix_rps10), c("sequence", "dada2_tax"))

if(!all(sample_cols_its == sample_cols_rps10)) {
  stop("Sample columns do not match between ITS and RPS10 dataframes!")
}

abundance <- rbind(
 asv_matrix_its[, c("sequence", "dada2_tax", sample_cols_its)],  # ITS data
  asv_matrix_rps10[, c("sequence", "dada2_tax", sample_cols_rps10)]  # RPS10 data
)

write.csv(abundance, "demulticoder/data/final_asv_abundance_matrix_combined_demulticoder.csv")

metadata_path <- file.path("demulticoder/data/metadata.csv")
metadata <- read_csv(metadata_path)
print(metadata)
```

### Load standard workflow outputs-ITS data
```{r load standard workflow outputs}
taxa_its<- read.table("standard_workflow/its/data/its_taxa.out", sep="\t", stringsAsFactors=F,header=T)

taxa_filt_its <- taxa_its %>%
  mutate(across(starts_with("tax."), ~ ifelse(get(str_replace(cur_column(), "tax.", "boot.")) < 0, NA, .))) %>%
  set_names(str_replace(names(.), "tax.", ""))

remove_prefix_its <- function(data) {
  data <- gsub("[a-z]__", "", data)  
  return(data)
}

taxa_filt_its[] <- lapply(taxa_filt_its, remove_prefix_its)

revise_species_its <- function(genus, species) {
  revise_species_its <- ifelse(is.na(species), NA, paste(genus, species, sep = "_"))
  return(revise_species_its)
}

taxa_filt_its$Species <- revise_species_its(taxa_filt_its$Genus, taxa_filt_its$Species)
taxa_its_df <- as.data.frame(taxa_filt_its)
taxa_its_df$sequence <- rownames(taxa_its_df)
taxa_its_df <- taxa_its_df[, c("sequence", setdiff(names(taxa_its_df), "sequence"))]
rownames(taxa_its_df) <- NULL

#Load ASV matrix now
seqtab_nochim_its <- read.table("standard_workflow/its/data/its_seqtab_nochim.out")
seqtab_nochim_its <- t(seqtab_nochim_its)
seqtab_nochim_its_df <- as.data.frame(seqtab_nochim_its)
seqtab_nochim_its_df$sequence <- rownames(seqtab_nochim_its_df)
seqtab_nochim_its_df <- seqtab_nochim_its_df[, c("sequence", setdiff(names(seqtab_nochim_its_df), "sequence"))]
rownames(seqtab_nochim_its_df) <- NULL


#Combine matrices
combined_df_its <- merge(seqtab_nochim_its_df, taxa_its_df, by = "sequence", all = TRUE)

taxonomic_cols <- c("Kingdom", "Phylum", "Class", "Order", 
                    "Family", "Genus", "Species")

bootstrap_cols <- c("boot.Kingdom", "boot.Phylum", "boot.Class", "boot.Order", 
                    "boot.Family", "boot.Genus", "boot.Species")

dada2_tax <- mapply(function(tax, boot, level) {
  paste(tax, boot, level, sep = "--")
}, combined_df_its[, taxonomic_cols], combined_df_its[, bootstrap_cols], 
   c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"))

combined_df_its$dada2_tax <- apply(dada2_tax, 1, paste, collapse = ";")

#Remove all other tax and boot columns
combined_df_its <- combined_df_its[, !(names(combined_df_its) %in% c(taxonomic_cols, bootstrap_cols))]

# Reorder columns to put 'taxa_bootstrap_combined' after 'sequence'
combined_df_its <- combined_df_its[, c("sequence", "dada2_tax", setdiff(names(combined_df_its), c("sequence", "dada2_tax")))]

#Add domain-100-Eukaryota; prefix before Kingdom
combined_df_its$dada2_tax <- gsub("Fungi", "Eukaryota--100--Domain;Fungi", combined_df_its$dada2_tax)

#Load metadata
samdf_its<- read.csv("standard_workflow/its/data/metadata_its.csv")

#Save file
write.csv(combined_df_its, "standard_workflow/its/data/its_abundance_matrix_combined_dada2workflow.csv")
```

### Load standard workflow outputs-rps10 data
```{r load standard workflow outputs}
taxa_rps10<- read.table("standard_workflow/rps10/data/rps10_taxa.out", sep="\t", stringsAsFactors=F,header=T)

taxa_filt_rps10 <- taxa_rps10 %>%
  mutate(across(starts_with("tax."), ~ ifelse(get(str_replace(cur_column(), "tax.", "boot.")) < 0, NA, .))) %>%
  set_names(str_replace(names(.), "tax.", ""))

remove_prefix_rps10 <- function(data) {
  data <- gsub("[a-z]__", "", data)  
  return(data)
}

taxa_filt_rps10[] <- lapply(taxa_filt_rps10, remove_prefix_rps10)

taxa_rps10_df <- as.data.frame(taxa_filt_rps10)
taxa_rps10_df$sequence <- rownames(taxa_rps10_df)
taxa_rps10_df <- taxa_rps10_df[, c("sequence", setdiff(names(taxa_rps10_df), "sequence"))]
rownames(taxa_rps10_df) <- NULL

#Load ASV matrix now
seqtab_nochim_rps10 <- read.table("standard_workflow/rps10/data/rps10_seqtab_nochim.out")
rownames(seqtab_nochim_rps10) <- gsub("__", "_", rownames(seqtab_nochim_rps10))

seqtab_nochim_rps10 <- t(seqtab_nochim_rps10)
seqtab_nochim_rps10_df <- as.data.frame(seqtab_nochim_rps10)
seqtab_nochim_rps10_df$sequence <- rownames(seqtab_nochim_rps10_df)
seqtab_nochim_rps10_df <- seqtab_nochim_rps10_df[, c("sequence", setdiff(names(seqtab_nochim_rps10_df), "sequence"))]
rownames(seqtab_nochim_rps10_df) <- NULL

#Combine matrices
combined_df_rps10 <- merge(seqtab_nochim_rps10_df, taxa_rps10_df, by = "sequence", all = TRUE)

taxonomic_cols <- c("Kingdom", "Phylum", "Class", "Order", 
                    "Family", "Genus", "Species")

bootstrap_cols <- c("boot.Kingdom", "boot.Phylum", "boot.Class", "boot.Order", 
                    "boot.Family", "boot.Genus", "boot.Species")

dada2_tax <- mapply(function(tax, boot, level) {
  paste(tax, boot, level, sep = "--")
}, combined_df_rps10[, taxonomic_cols], combined_df_rps10[, bootstrap_cols], 
   c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"))

combined_df_rps10$dada2_tax <- apply(dada2_tax, 1, paste, collapse = ";")

#Remove all other tax and boot columns
combined_df_rps10 <- combined_df_rps10[, !(names(combined_df_rps10) %in% c(taxonomic_cols, bootstrap_cols))]

# Reorder columns to put 'taxa_bootstrap_combined' after 'sequence'
combined_df_rps10 <- combined_df_rps10[, c("sequence", "dada2_tax", setdiff(names(combined_df_rps10), c("sequence", "dada2_tax")))]

#Add domain-100-Eukaryota; prefix before Kingdom
combined_df_rps10$dada2_tax <- gsub("Stramenopila", "Eukaryota--100--Domain;Stramenopila", combined_df_rps10$dada2_tax)

#Load metadata
samdf_rps10<- read.csv("standard_workflow/rps10/data/metadata_rps10.csv")

#Save file
write.csv(combined_df_rps10, "standard_workflow/rps10/data/rps10_abundance_matrix_combined_dada2workflow.csv")

#Load DADA2 metadata file
samdf_rps10<- read.csv("standard_workflow/rps10/data/metadata_rps10.csv")
```
```

### Generate summary of read counts throughout the workflow for demlticoder specfific analysis
```{r load demulticoder outputs}
#Fourth, let's track reads
track_reads_demulticoder_its<-read.csv("demulticoder/data/track_reads_its.csv", row.names = 1)
track_reads_demulticoder_rps10<-read.csv("demulticoder/data/track_reads_rps10.csv", row.names = 1)
```

### Generate summary of read counts throughout the workflow for demlticoder specfific analysis
```{r load demulticoder outputs}

```

### Track reads throughout demulticoder workflows

### Track reasd throughout standard workflows
```{r}
sessioninfo::session_info()
```