---
title: "A comparison of standard workflow vs. demulticoder analysis outputs"
output: 
  rmarkdown::html_vignette:
    fig_path: "man/figures"
vignette: >
  %\VignetteIndexEntry{Multiplexed and pooled barcode examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(phyloseq); packageVersion("phyloseq")
library(ggplot2); packageVersion("ggplot2")
library(readr); packageVersion("readr")
library(tidyr); packageVersion("tidyr")
library(purrr); packageVersion("purrr")
library(furrr); packageVersion("furrr")
library(dplyr); packageVersion("dplyr")
library(stringr); packageVersion("stringr")
library(forcats); packageVersion("forcats")
library(metacoder); packageVersion("metacoder")
library(data.table); packageVersion("data.table")
library(decontam); packageVersion("decontam")
library(Biostrings); packageVersion("Biostrings")
library(magick); packageVersion("magick")
library(vegan); packageVersion("vegan")
library(pdftools);packageVersion("pdftools")
library(vegan); packageVersion("vegan")
library(grid)
library(gridExtra)
```

## More in-depth comparison of standard workflow outputs vs. demulticoder analysis outputs for a multiplexed and pooled pooled barcode dataset

This analysis accompanies the rhododendron rhizobiome analyses described in the associated manuscript. 

Briefly, TODO

### Input metadata and primerinfo_params files are in data folder
```{r setup, include=FALSE}
#TODO-at some point switch to loading github repo version
devtools::install_github("grunwaldlab/demulticoder", force=TRUE)
library("demulticoder")
```

### Load demulticoder outputs
```{r load demulticoder outputs}
seed <- 1
set.seed(seed)

asv_matrix_rps10<-read.csv("~/benchmark_demulticoder/demulticoder/data/final_asv_abundance_matrix_rps10.csv")
asv_matrix_rps10$dada2_tax <- asv_matrix_rps10$dada2_tax <- gsub("Stramenopila", "Eukaryota--100--Domain;Stramenopila", asv_matrix_rps10$dada2_tax)
asv_matrix_rps10 <- asv_matrix_rps10[, -1]
colnames(asv_matrix_rps10)[3:ncol(asv_matrix_rps10)] <- gsub("_rps10$", "", colnames(asv_matrix_rps10)[3:ncol(asv_matrix_rps10)])

asv_matrix_its<-read.csv("~/benchmark_demulticoder/demulticoder/data/final_asv_abundance_matrix_its.csv")
asv_matrix_its$dada2_tax <- gsub("Fungi", "Eukaryota--100--Domain;Fungi", asv_matrix_its$dada2_tax)
asv_matrix_its <- asv_matrix_its[, -1]
colnames(asv_matrix_its)[3:ncol(asv_matrix_rps10)] <- gsub("_its$", "", colnames(asv_matrix_its)[3:ncol(asv_matrix_its)])


#Let's combine these matrices
#For easier analysis, we previously combined the two matrices, and appended domain info to each one so we can make one heat tree for combined dataset
sample_cols_its <- setdiff(names(asv_matrix_its), c("sequence", "dada2_tax"))
sample_cols_rps10 <- setdiff(names(asv_matrix_rps10), c("sequence", "dada2_tax"))

if(!all(sample_cols_its == sample_cols_rps10)) {
  stop("Sample columns do not match between ITS and RPS10 dataframes!")
}

abundance <- rbind(
 asv_matrix_its[, c("sequence", "dada2_tax", sample_cols_its)],  # ITS data
  asv_matrix_rps10[, c("sequence", "dada2_tax", sample_cols_rps10)]  # RPS10 data
)

abundance_with_inf <- read.csv("~/benchmark_demulticoder/demulticoder/data/abundance_with_infection_data.csv") 

metadata_path <- file.path("~/benchmark_demulticoder/demulticoder/data/metadata_with_infection_data.csv")
metadata <- read_csv(metadata_path)
print(metadata)
```

### Load standard workflow outputs-ITS data
```{r load standard workflow outputs}
taxa_its<- read.table("~/benchmark_demulticoder/standard_workflow/its/data/its_taxa.out", sep="\t", stringsAsFactors=F,header=T)

taxa_filt_its <- taxa_its %>%
  mutate(across(starts_with("tax."), ~ ifelse(get(str_replace(cur_column(), "tax.", "boot.")) < 0, NA, .))) %>%
  set_names(str_replace(names(.), "tax.", ""))

remove_prefix_its <- function(data) {
  data <- gsub("[a-z]__", "", data)  
  return(data)
}

taxa_filt_its[] <- lapply(taxa_filt_its, remove_prefix_its)

revise_species_its <- function(genus, species) {
  revise_species_its <- ifelse(is.na(species), NA, paste(genus, species, sep = "_"))
  return(revise_species_its)
}

taxa_filt_its$Species <- revise_species_its(taxa_filt_its$Genus, taxa_filt_its$Species)
taxa_its_df <- as.data.frame(taxa_filt_its)
taxa_its_df$sequence <- rownames(taxa_its_df)
taxa_its_df <- taxa_its_df[, c("sequence", setdiff(names(taxa_its_df), "sequence"))]
rownames(taxa_its_df) <- NULL

#Load ASV matrix now
seqtab_nochim_its <- read.table("~/benchmark_demulticoder/standard_workflow/its/data/its_seqtab_nochim.out")
seqtab_nochim_its <- t(seqtab_nochim_its)
seqtab_nochim_its_df <- as.data.frame(seqtab_nochim_its)
seqtab_nochim_its_df$sequence <- rownames(seqtab_nochim_its_df)
seqtab_nochim_its_df <- seqtab_nochim_its_df[, c("sequence", setdiff(names(seqtab_nochim_its_df), "sequence"))]
rownames(seqtab_nochim_its_df) <- NULL


#Combine matrices
combined_df_its <- merge(seqtab_nochim_its_df, taxa_its_df, by = "sequence", all = TRUE)

taxonomic_cols <- c("Kingdom", "Phylum", "Class", "Order", 
                    "Family", "Genus", "Species")

bootstrap_cols <- c("boot.Kingdom", "boot.Phylum", "boot.Class", "boot.Order", 
                    "boot.Family", "boot.Genus", "boot.Species")

dada2_tax <- mapply(function(tax, boot, level) {
  paste(tax, boot, level, sep = "--")
}, combined_df_its[, taxonomic_cols], combined_df_its[, bootstrap_cols], 
   c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"))

combined_df_its$dada2_tax <- apply(dada2_tax, 1, paste, collapse = ";")

#Remove all other tax and boot columns
combined_df_its <- combined_df_its[, !(names(combined_df_its) %in% c(taxonomic_cols, bootstrap_cols))]

# Reorder columns to put 'taxa_bootstrap_combined' after 'sequence'
combined_df_its <- combined_df_its[, c("sequence", "dada2_tax", setdiff(names(combined_df_its), c("sequence", "dada2_tax")))]

#Add domain-100-Eukaryota; prefix before Kingdom
combined_df_its$dada2_tax <- gsub("Fungi", "Eukaryota--100--Domain;Fungi", combined_df_its$dada2_tax)

#Load metadata
samdf_its<- read.csv("~/benchmark_demulticoder/standard_workflow/its/data/metadata_its.csv")
```

### Load standard workflow outputs-rps10 data
```{r load standard workflow outputs}
taxa_rps10<- read.table("~/benchmark_demulticoder/standard_workflow/rps10/data/rps10_taxa.out", sep="\t", stringsAsFactors=F,header=T)

taxa_filt_rps10 <- taxa_rps10 %>%
  mutate(across(starts_with("tax."), ~ ifelse(get(str_replace(cur_column(), "tax.", "boot.")) < 0, NA, .))) %>%
  set_names(str_replace(names(.), "tax.", ""))

remove_prefix_rps10 <- function(data) {
  data <- gsub("[a-z]__", "", data)  
  return(data)
}

taxa_filt_rps10[] <- lapply(taxa_filt_rps10, remove_prefix_rps10)

taxa_rps10_df <- as.data.frame(taxa_filt_rps10)
taxa_rps10_df$sequence <- rownames(taxa_rps10_df)
taxa_rps10_df <- taxa_rps10_df[, c("sequence", setdiff(names(taxa_rps10_df), "sequence"))]
rownames(taxa_rps10_df) <- NULL

#Load ASV matrix now
seqtab_nochim_rps10 <- read.table("~/benchmark_demulticoder/standard_workflow/rps10/data/rps10_seqtab_nochim.out")
rownames(seqtab_nochim_rps10) <- gsub("__", "_", rownames(seqtab_nochim_rps10))

seqtab_nochim_rps10 <- t(seqtab_nochim_rps10)
seqtab_nochim_rps10_df <- as.data.frame(seqtab_nochim_rps10)
seqtab_nochim_rps10_df$sequence <- rownames(seqtab_nochim_rps10_df)
seqtab_nochim_rps10_df <- seqtab_nochim_rps10_df[, c("sequence", setdiff(names(seqtab_nochim_rps10_df), "sequence"))]
rownames(seqtab_nochim_rps10_df) <- NULL

#Combine matrices
combined_df_rps10 <- merge(seqtab_nochim_rps10_df, taxa_rps10_df, by = "sequence", all = TRUE)

taxonomic_cols <- c("Kingdom", "Phylum", "Class", "Order", 
                    "Family", "Genus", "Species")

bootstrap_cols <- c("boot.Kingdom", "boot.Phylum", "boot.Class", "boot.Order", 
                    "boot.Family", "boot.Genus", "boot.Species")

dada2_tax <- mapply(function(tax, boot, level) {
  paste(tax, boot, level, sep = "--")
}, combined_df_rps10[, taxonomic_cols], combined_df_rps10[, bootstrap_cols], 
   c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"))

combined_df_rps10$dada2_tax <- apply(dada2_tax, 1, paste, collapse = ";")

#Remove all other tax and boot columns
combined_df_rps10 <- combined_df_rps10[, !(names(combined_df_rps10) %in% c(taxonomic_cols, bootstrap_cols))]

# Reorder columns to put 'taxa_bootstrap_combined' after 'sequence'
combined_df_rps10 <- combined_df_rps10[, c("sequence", "dada2_tax", setdiff(names(combined_df_rps10), c("sequence", "dada2_tax")))]

#Add domain-100-Eukaryota; prefix before Kingdom
combined_df_rps10$dada2_tax <- gsub("Stramenopila", "Eukaryota--100--Domain;Stramenopila", combined_df_rps10$dada2_tax)

#Load metadata
samdf_rps10<- read.csv("~/benchmark_demulticoder/standard_workflow/rps10/data/metadata_rps10.csv")

#Load DADA2 metadata file
samdf_rps10<- read.csv("~/benchmark_demulticoder/standard_workflow/rps10/data/metadata_rps10.csv")
```

### Finally Let's first combine rps10 and its data into fully combined matrices
```{r combine dataframes standard workflows}
sample_cols_its <- setdiff(names(combined_df_its), c("sequence", "dada2_tax"))
sample_cols_rps10 <- setdiff(names(combined_df_rps10), c("sequence", "dada2_tax"))

if(!all(sample_cols_its == sample_cols_rps10)) {
  stop("Sample columns do not match between ITS and RPS10 dataframes!")
}

abundance_sw <- rbind(
  combined_df_its[, c("sequence", "dada2_tax", sample_cols_its)],  # ITS data
  combined_df_rps10[, c("sequence", "dada2_tax", sample_cols_rps10)]  # RPS10 data
)

abundance_sw_inf <- read.csv("~/benchmark_demulticoder/standard_workflow/combined_data/metadata_with_infection_data.csv")

#Let's also load metadata file-consistent with demulticoder analysis
metadata_path_sf <- file.path("~/benchmark_demulticoder/standard_workflow/combined_data/metadata_with_infection_data.csv")
metadata_sf <- read_csv(metadata_path_sf)
print(metadata_sf)
```

### Finally Let's first combine rps10 and its data into fully combined matrices
```{r combine dataframes}
sample_cols_its <- setdiff(names(combined_df_its), c("sequence", "dada2_tax"))
sample_cols_rps10 <- setdiff(names(combined_df_rps10), c("sequence", "dada2_tax"))

if(!all(sample_cols_its == sample_cols_rps10)) {
  stop("Sample columns do not match between ITS and RPS10 dataframes!")
}

abundance <- rbind(
  combined_df_its[, c("sequence", "dada2_tax", sample_cols_its)],  # ITS data
  combined_df_rps10[, c("sequence", "dada2_tax", sample_cols_rps10)]  # RPS10 data
)


#Let's also load metadata file-consistent with demulticoder analysis
metadata_path <- file.path("~/benchmark_demulticoder/standard_workflow/combined_data/metadata_with_infection_data.csv")
metadata <- read_csv(metadata_path)
print(metadata)
```

### Generate summary of read counts throughout the workflow for demlticoder specific analysis
```{r get demulticoder read stats}
#Fourth, let's track reads
track_reads_demulticoder_its<-read.csv("~/benchmark_demulticoder/demulticoder/data/track_reads_its.csv", row.names = 1)
track_reads_demulticoder_rps10<-read.csv("~/benchmark_demulticoder/demulticoder/data/track_reads_rps10.csv", row.names = 1)
track_reads_sw_its<-read.csv("~/benchmark_demulticoder/standard_workflow/its/data/its_trackreads.out", row.names = 1)
track_reads_standard_wf_its <- read.table("~/benchmark_demulticoder/standard_workflow/its/data/its_trackreads.out", sep="\t", stringsAsFactors = F, header=T)
track_reads_standard_wf_rps10 <- read.table("~/benchmark_demulticoder/standard_workflow/rps10/data/rps10_trackreads.out", sep="\t", stringsAsFactors = F, header=T)
```

### Now, let's compare read counts across workflows for rps10 amplicons
```{r compare reads-rps10}
rps10_summary_reads_demulticoder<-summary(track_reads_demulticoder_rps10)
print(rps10_summary_reads_demulticoder)

rps10_summary_reads_sw<-summary(track_reads_standard_wf_rps10)
print(rps10_summary_reads_sw)
```

### Now, let's compare read counts across workflows for ITS amplicons
```{r compare reads-rps10}
its_summary_reads_demulticoder<-summary(track_reads_demulticoder_its)
print(its_summary_reads_demulticoder)

its_summary_reads_sw<-summary(track_reads_standard_wf_its)
print(its_summary_reads_sw)
```

### Let's convert our two matrices to taxmap objects
Let's convert our new matrix and associated metadata table to a taxmap object for demulticoder analysis
```{r, Convert matrix created with demulticoder workflow to taxmap obj}
metadata <- filter(metadata, ! is_ambiguous)
abundance_with_inf <- filter(abundance_with_inf, ! is_inoculum)

#Filter out mock community samples and also any mock community samples as well
metadata <- filter(metadata, sample_type != "Mock community" & sample_type != "Negative control")

obj <- parse_tax_data(abundance, class_cols = 'dada2_tax', class_sep = ';',
                      class_regex = '^(.+)--(.+)--(.+)$',
                      class_key = c(taxon = 'taxon_name', boot = 'info', rank = 'taxon_rank'))
names(obj$data) <- c('abund', 'score')
obj <- transmute_obs(obj, 'score', sequence = sequence[input_index], boot = boot, rank = rank)


# For diversity calculations, we'll use proportions of read depth
# We'll set any proportion to 0 that is less than the inverse of the read count of the non-control sample with the fewest reads.
# This should account for unequal sample read depth without the randomness of rarefaction.

metadata$raw_count <- colSums(obj$data$abund[, metadata$sample_name])
lowest_count <- min(metadata$raw_count[! is.na(metadata$organism)])
lowest_count
obj$data$prop <- calc_obs_props(obj, data = 'abund', cols = metadata$sample_name)
obj$data$prop <- zero_low_counts(obj, data = 'prop', min_count = 1 / lowest_count, cols = metadata$sample_name)
obj$data$prop

obj$data$prop[metadata$sample_name] <- map(metadata$sample_name, function(id) {
  out <- obj$data$prop[[id]]
  out[is.na(out) | is.nan(out)] <- 0
  out
})

obj$data$prop

#Let's modify the metadata sheet
metadata <- metadata %>%
  mutate(organism = fct_relevel(ordered(organism), "Control", "Cin", "Plu", "Cry"))
```

Let's convert our new matrix and associated metadata table to a taxmap object for output of standard workflow analysis
```{r, Convert matrix created with demulticoder workflow to taxmap obj}
metadata_sf <- filter(metadata, ! is_ambiguous)
abundance_sw_inf <- filter(abundance_with_inf, ! is_inoculum)

#Filter out mock community samples and also any mock community samples as well
metadata_sf <- filter(metadata, sample_type != "Mock community" & sample_type != "Negative control")

obj_sf <- parse_tax_data(abundance_sw_inf, class_cols = 'dada2_tax', class_sep = ';',
                      class_regex = '^(.+)--(.+)--(.+)$',
                      class_key = c(taxon = 'taxon_name', boot = 'info', rank = 'taxon_rank'))
names(obj_sf$data) <- c('abund', 'score')
obj_sf <- transmute_obs(obj_sf, 'score', sequence = sequence[input_index], boot = boot, rank = rank)


# For diversity calculations, we'll use proportions of read depth
# We'll set any proportion to 0 that is less than the inverse of the read count of the non-control sample with the fewest reads.
# This should account for unequal sample read depth without the randomness of rarefaction.

metadata_sf$raw_count <- colSums(obj_sf$data$abund[, metadata_sf$sample_name])
lowest_count <- min(metadata_sf$raw_count[! is.na(metadata_sf$organism)])
lowest_count
obj_sf$data$prop <- calc_obs_props(obj_sf, data = 'abund', cols = metadata_sf$sample_name)
obj_sf$data$prop <- zero_low_counts(obj_sf, data = 'prop', min_count = 1 / lowest_count, cols = metadata_sf$sample_name)
obj_sf$data$prop

obj_sf$data$prop[metadata_sf$sample_name] <- map(metadata_sf$sample_name, function(id) {
  out <- obj_sf$data$prop[[id]]
  out[is.na(out) | is.nan(out)] <- 0
  out
})

obj_sf$data$prop

#Let's modify the metadata_sf sheet
metadata_sf <- metadata_sf %>%
  mutate(organism = fct_relevel(ordered(organism), "Control", "Cin", "Plu", "Cry"))
```

### Let's examine alpha diversity within pots
Demulticoder outputs
```{r, alpha diversity demulticoder}
abund_table <- obj$data$abund[metadata$sample_name]
metadata$richness <- vegan::specnumber(abund_table, MARGIN = 2)
metadata$shannon <- vegan::diversity(abund_table, MARGIN = 2, index = "shannon")
metadata$invsimpson <- vegan::diversity(abund_table, MARGIN = 2, index = "invsimpson")

plotted_factors <- c('Organism' = 'organism', 'Flooded' = 'flooded', 'Pathogen Concentration' = 'path_conc', 'Trial' = 'experiment')

# Reformat data for plotting
alpha_plot_data <- plotted_factors %>%
  map2_dfr(names(plotted_factors), 
           function(factor, factor_name) {
             out <- metadata
             out$factor <- factor_name
             out$value <- as.character(metadata[[factor]])
             return(out)
           }) %>%
  mutate(path_conc = factor(path_conc, 
                            levels = sort(unique(path_conc)), 
                            labels = paste(sort(unique(path_conc)), 'CFU/g'), 
                            ordered = TRUE)) %>%
  filter(sample_type == 'Sample') %>%
  select(sample_name, factor, value, invsimpson) %>%
  tidyr::gather(key = "index", value = "diversity", -sample_name, -factor, -value) %>%
  mutate(value = forcats::fct_relevel(ordered(value), "Control", "Cin", "Plu", "Cry"))

# ANOVA and Tukey's HSD
anova_and_hsd <- function(x) {
  anova_result <- aov(diversity ~ value, x)
  tukey_result <- agricolae::HSD.test(anova_result, "value", group = TRUE)
  group_data <- tukey_result$groups[order(rownames(tukey_result$groups)),]
  group_key <- setNames(group_data$groups, rownames(group_data))
  group_key[as.character(x$value)]
}
alpha_plot_data$group <- unlist(map(split(alpha_plot_data, alpha_plot_data$factor)[unique(alpha_plot_data$factor)], anova_and_hsd))


alpha_subplot <- ggplot(alpha_plot_data, aes(x = value, y = diversity)) +
  geom_boxplot() +
  geom_text(aes(x = value,
                y = max(diversity) + 2,
                label = group),
            col = 'black',
            size = 5) +
  facet_grid( ~ factor, scales = "free") +
  labs(x = NULL, y = 'Diversity (Inverse Simpson)') +
  guides(color = "none") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom")
```

Standard workflow
```{r, alpha diversity demulticoder}
abund_table <- obj$data$abund[metadata$sample_name]
metadata$richness <- vegan::specnumber(abund_table, MARGIN = 2)
metadata$shannon <- vegan::diversity(abund_table, MARGIN = 2, index = "shannon")
metadata$invsimpson <- vegan::diversity(abund_table, MARGIN = 2, index = "invsimpson")
write.csv(metadata, file.path("standard_workflow", "combined_results/alpha_diversity.csv"))
print(metadata)

plotted_factors <- c('Organism' = 'organism', 'Flooded' = 'flooded', 'Pathogen Concentration' = 'path_conc', 'Trial' = 'experiment')

# Reformat data for plotting
alpha_plot_data <- plotted_factors %>%
  map2_dfr(names(plotted_factors), 
           function(factor, factor_name) {
             out <- metadata
             out$factor <- factor_name
             out$value <- as.character(metadata[[factor]])
             return(out)
           }) %>%
  mutate(path_conc = factor(path_conc, 
                            levels = sort(unique(path_conc)), 
                            labels = paste(sort(unique(path_conc)), 'CFU/g'), 
                            ordered = TRUE)) %>%
  filter(sample_type == 'Sample') %>%
  select(sample_name, factor, value, invsimpson) %>%
  tidyr::gather(key = "index", value = "diversity", -sample_name, -factor, -value) %>%
  mutate(value = forcats::fct_relevel(ordered(value), "Control", "Cin", "Plu", "Cry"))

# ANOVA and Tukey's HSD
anova_and_hsd <- function(x) {
  anova_result <- aov(diversity ~ value, x)
  tukey_result <- agricolae::HSD.test(anova_result, "value", group = TRUE)
  group_data <- tukey_result$groups[order(rownames(tukey_result$groups)),]
  group_key <- setNames(group_data$groups, rownames(group_data))
  group_key[as.character(x$value)]
}
alpha_plot_data$group <- unlist(map(split(alpha_plot_data, alpha_plot_data$factor)[unique(alpha_plot_data$factor)], anova_and_hsd))


alpha_subplot <- ggplot(alpha_plot_data, aes(x = value, y = diversity)) +
  geom_boxplot() +
  geom_text(aes(x = value,
                y = max(diversity) + 2,
                label = group),
            col = 'black',
            size = 5) +
  facet_grid( ~ factor, scales = "free") +
  labs(x = NULL, y = 'Diversity (Inverse Simpson)') +
  guides(color = "none") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom")
```

### Let's examine beta diversity within pots
Demulticoder outputs
```{r, beta diversity demulticoder}
set.seed(1)
prob_table <- obj$data$prop[metadata$sample_name]
nmds_plot_data <- function(prob_table) {
  metadata <- metadata[metadata$sample_name %in% colnames(prob_table), ]
  set.seed(1)
  nmds_results <- vegan::metaMDS(t(prob_table), trymax = 1000, k = 2, trace = 0)
  nmds_data <- nmds_results$points %>%
    as_tibble() %>%
    bind_cols(metadata)
  names(nmds_data)[1:2] <- paste0("NMDS", 1:2)
  return(nmds_data)
}

nmds_data <- nmds_plot_data(prob_table[! is.na(metadata$organism) & metadata$valid_inoc])

nmds_factors <- c(Flooded = 'flooded', Organism = 'organism', 'Pathogen CFU/g' = 'path_conc', 'Trial' = 'experiment')

make_one_plot <- function(factor, name) {
  nmds_data %>%
    mutate(factor = as.character(nmds_data[[factor]]),
           NMDS1 = scales::rescale(NMDS1),
           NMDS2 = scales::rescale(NMDS2)) %>%
    mutate(factor = fct_relevel(ordered(factor), "Control", "Cin", "Plu", "Cry")) %>%
    ggplot(aes_string(x = "NMDS1", y = "NMDS2", color = "factor", label = "sample_name")) +
    # geom_label(size = 2) +
    geom_point() +
    coord_fixed() +
    viridis::scale_color_viridis(discrete = TRUE, end = .9) +
    labs(color = NULL, x = NULL, y = NULL) +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text = element_blank(),
          axis.title = element_text(size = 7), 
          axis.ticks = element_blank(),
          plot.margin = unit(rep(0.04, 4), "cm"),
          # panel.background = element_rect(fill = 'transparent', colour = NA),
          # plot.background = element_rect(fill = "white", colour = NA),
          legend.position = "bottom",
          legend.text = element_text(size = 10),
          legend.key.height = unit(0.5, 'cm'),
          legend.key.width = unit(0.2, 'cm'))
}

nmds_subplots <- map2(nmds_factors, names(nmds_factors), make_one_plot)
nmds_plot <- ggpubr::ggarrange(plotlist = c(list(ggplot() + theme_void()), nmds_subplots),
                       nrow = 1, widths = c(0.15, 1, 1, 1, 1))

ggsave(nmds_plot, path = "standard_workflow/combined_figures", filename = "nmds.pdf", 
       width = 7, height = 8)
print(nmds_plot)

combined_div_plot <- ggpubr::ggarrange(alpha_subplot, nmds_plot, ncol = 1, labels = c('A', 'B'),
                               heights = c(1, 1))
combined_div_plot

ggsave(combined_div_plot, path = "standard_workflow/combined_figures", filename = "diversity.pdf", 
       width = 10, height = 6, bg = "#FFFFFF")

ggsave(combined_div_plot, path = "standard_workflow/combined_figures", filename = "diversity.svg", 
       width = 10, height = 6, bg = "#FFFFFF")
```

Standard workflow
```{r, beta diversity demulticoder}
set.seed(1)
prob_table <- obj$data$prop[metadata$sample_name]
nmds_plot_data <- function(prob_table) {
  metadata <- metadata[metadata$sample_name %in% colnames(prob_table), ]
  set.seed(1)
  nmds_results <- vegan::metaMDS(t(prob_table), trymax = 1000, k = 2, trace = 0)
  nmds_data <- nmds_results$points %>%
    as_tibble() %>%
    bind_cols(metadata)
  names(nmds_data)[1:2] <- paste0("NMDS", 1:2)
  return(nmds_data)
}

nmds_data <- nmds_plot_data(prob_table[! is.na(metadata$organism) & metadata$valid_inoc])

nmds_factors <- c(Flooded = 'flooded', Organism = 'organism', 'Pathogen CFU/g' = 'path_conc', 'Trial' = 'experiment')

make_one_plot <- function(factor, name) {
  nmds_data %>%
    mutate(factor = as.character(nmds_data[[factor]]),
           NMDS1 = scales::rescale(NMDS1),
           NMDS2 = scales::rescale(NMDS2)) %>%
    mutate(factor = fct_relevel(ordered(factor), "Control", "Cin", "Plu", "Cry")) %>%
    ggplot(aes_string(x = "NMDS1", y = "NMDS2", color = "factor", label = "sample_name")) +
    # geom_label(size = 2) +
    geom_point() +
    coord_fixed() +
    viridis::scale_color_viridis(discrete = TRUE, end = .9) +
    labs(color = NULL, x = NULL, y = NULL) +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text = element_blank(),
          axis.title = element_text(size = 7), 
          axis.ticks = element_blank(),
          plot.margin = unit(rep(0.04, 4), "cm"),
          # panel.background = element_rect(fill = 'transparent', colour = NA),
          # plot.background = element_rect(fill = "white", colour = NA),
          legend.position = "bottom",
          legend.text = element_text(size = 10),
          legend.key.height = unit(0.5, 'cm'),
          legend.key.width = unit(0.2, 'cm'))
}

nmds_subplots <- map2(nmds_factors, names(nmds_factors), make_one_plot)
nmds_plot <- ggpubr::ggarrange(plotlist = c(list(ggplot() + theme_void()), nmds_subplots),
                       nrow = 1, widths = c(0.15, 1, 1, 1, 1))

ggsave(nmds_plot, path = "standard_workflow/combined_figures", filename = "nmds.pdf", 
       width = 7, height = 8)
print(nmds_plot)

combined_div_plot <- ggpubr::ggarrange(alpha_subplot, nmds_plot, ncol = 1, labels = c('A', 'B'),
                               heights = c(1, 1))
combined_div_plot

ggsave(combined_div_plot, path = "standard_workflow/combined_figures", filename = "diversity.pdf", 
       width = 10, height = 6, bg = "#FFFFFF")

ggsave(combined_div_plot, path = "standard_workflow/combined_figures", filename = "diversity.svg", 
       width = 10, height = 6, bg = "#FFFFFF")
```

### Let's do a quick differential abundance analysis with DESeq2
Demulticoder outputs
```{r, alpha diversity demulticoder}
```

Standard workflow
```{r, alpha diversity demulticoder}
```

### Let's examine community composition more broadly
Demulticoder outputs
```{r, alpha diversity demulticoder}
```

Standard workflow
```{r, alpha diversity demulticoder}
```

```{r}
sessioninfo::session_info()
```